---
title: "Insead Big Data Class- Speed dating"
output: html_document
Authors: Judith Finegold, Kacper Ostrowski, Ulf Nore, Maciej Piotrowicz
---

```{r setuplibraries, echo=FALSE, message=FALSE}
suppressWarnings(source("./AnalyticsLibraries/library.R"))
# Package options
suppressWarnings(ggthemr('fresh'))  # ggplot theme
opts_knit$set(progress=FALSE, verbose=FALSE)
opts_chunk$set(echo=FALSE, fig.align="center", fig.width=10, fig.height=6.35, results="asis")
options(knitr.kable.NA = '')
```




`r rm(list=ls(all=TRUE)) `

```{r echo=FALSE, eval=FALSE, tidy=TRUE}
##setwd("/GroupProject") COMMENT:We don't need to set a new directory because the project opens in the right directory anyway
list.files()
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(out.width='900px', dpi=200)
#opts_chunk$set(dev = 'pdf')

library(dplyr)
library(ggplot2)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(foreign)
library(broom)
library(corrgram)
library(effects)
library(car)
library(caret)
library(gridExtra)
library(knitr)
library(tidyr)
```

```{r echo=FALSE, eval=TRUE, tidy=TRUE}
sdd <- read.csv("./Data/Testdata.csv", sep=",", dec="." ,header=TRUE)
#sdd=data.matrix(sdd)
```


# The Data

First we load the data to use (see the raw .Rmd file to change the data file as needed):

```{r setupdata1E, echo=TRUE, tidy=TRUE}
# Please ENTER the name of the file with the data used. The file should be a .csv with one row per observation (e.g. person) and one column per attribute. Do not add .csv at the end, make sure the data are numeric.
datafile_name = "./Data/Testdata.csv"

# Please enter the minimum number below which you would like not to print - this makes the readability of the tables easier. Default values are either 10e6 (to print everything) or 0.5. Try both to see the difference.
MIN_VALUE = 0.5

# Please enter the maximum number of observations to show in the report and slides. 
# DEFAULT is 10. If the number is large the report may be slow.
max_data_report = 10
```

```{r}
d <- read.csv(datafile_name)
d <- data.matrix(d) 
ProjectData_INITIAL <- d

```

<hr>\clearpage

# Part 1: Key Customer Characteristics


```{r setupfactor, echo=TRUE, tidy=TRUE}
# Please ENTER then original raw attributes to use. 
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
factor_attributes_used = c(6:15)

# Please ENTER the selection criterions for the factors to use. 
# Choices: "eigenvalue", "variance", "manual"
factor_selectionciterion = "manual"

# Please ENTER the desired minumum variance explained 
# (Only used in case "variance" is the factor selection criterion used). 
minimum_variance_explained = 65  # between 1 and 100

# Please ENTER the number of factors to use 
# (Only used in case "manual" is the factor selection criterion used).
manual_numb_factors_used = 7

# Please ENTER the rotation eventually used (e.g. "none", "varimax", "quatimax", "promax", "oblimin", "simplimax", and "cluster" - see help(principal)). Default is "varimax"
rotation_used = "varimax"

```

```{r}
factor_attributes_used <- intersect(factor_attributes_used, 1:ncol(d))
ProjectDataFactor <- d[,factor_attributes_used]
ProjectDataFactor <- d <- data.matrix(ProjectDataFactor)
```

## Steps 1-2: Check the Data 


Start by some basic visual exploration of, say, a few data:

```{r}
rownames(ProjectDataFactor) <- paste0("Date ", sprintf("%02i", 1:nrow(ProjectDataFactor)))
iprint.df(t(head(round(ProjectDataFactor, 2), max_data_report)))
```

The data we use here have the following descriptive statistics: 

```{r}
iprint.df(round(my_summary(ProjectDataFactor), 2))
```

##should work if data set is clean and no empty rows


## Step 3: Check Correlations

This is the correlation matrix of the customer responses to the `r ncol(ProjectDataFactor)` attitude questions - which are the only questions that we will use for the segmentation (see the case):

```{r}
thecor = round(cor(ProjectDataFactor),2)
iprint.df(round(thecor,2), scale=TRUE)
```


## Step 4: Choose number of factors


```{r}
# Here is how the `principal` function is used 
UnRotated_Results<-principal(ProjectDataFactor, nfactors=ncol(ProjectDataFactor), rotate="none",score=TRUE)
UnRotated_Factors<-round(UnRotated_Results$loadings,2)
UnRotated_Factors<-as.data.frame(unclass(UnRotated_Factors))
colnames(UnRotated_Factors)<-paste("Comp",1:ncol(UnRotated_Factors),sep="")
```

```{r}
# Here is how we use the `PCA` function 
Variance_Explained_Table_results<-PCA(ProjectDataFactor, graph=FALSE)
Variance_Explained_Table<-Variance_Explained_Table_results$eig
Variance_Explained_Table_copy<-Variance_Explained_Table

rownames(Variance_Explained_Table) <- paste("Component", 1:nrow(Variance_Explained_Table), sep=" ")
colnames(Variance_Explained_Table) <- c("Eigenvalue", "Pct of explained variance", "Cumulative pct of explained variance")
```

Let's look at the **variance explained** as well as the **eigenvalues** (see session readings):

```{r}
iprint.df(round(Variance_Explained_Table, 2))
```

```{r}
eigenvalues  <- Variance_Explained_Table[, "Eigenvalue"]
df           <- cbind(as.data.frame(eigenvalues), c(1:length(eigenvalues)), rep(1, length(eigenvalues)))
colnames(df) <- c("eigenvalues", "components", "abline")
iplot.df(melt(df, id="components"))
```


## Step 5: Interpret the factors

Let's now see how the "top factors" look like. 

```{r}
if (factor_selectionciterion == "eigenvalue")
  factors_selected = sum(Variance_Explained_Table_copy[,1] >= 1)
if (factor_selectionciterion == "variance")
  factors_selected = 1:head(which(Variance_Explained_Table_copy[,"cumulative percentage of variance"]>= minimum_variance_explained),1)
if (factor_selectionciterion == "manual")
  factors_selected = manual_numb_factors_used
```


To better visualize them, we will use what is called a "rotation". There are many rotations methods. In this case we selected the `r rotation_used` rotation. For our data, the `r factors_selected` selected factors look as follows after this rotation: 

```{r}
Rotated_Results<-principal(ProjectDataFactor, nfactors=max(factors_selected), rotate=rotation_used,score=TRUE)
Rotated_Factors<-round(Rotated_Results$loadings,2)
Rotated_Factors<-as.data.frame(unclass(Rotated_Factors))
colnames(Rotated_Factors)<-paste("Comp.",1:ncol(Rotated_Factors),sep="")

sorted_rows <- sort(Rotated_Factors[,1], decreasing = TRUE, index.return = TRUE)$ix
Rotated_Factors <- Rotated_Factors[sorted_rows,]

iprint.df(Rotated_Factors, scale=TRUE)
```

To better visualize and interpret the factors we often "suppress" loadings with small values, e.g. with absolute values smaller than 0.5. In this case our factors look as follows after suppressing the small numbers:

```{r}
Rotated_Factors_thres <- Rotated_Factors
Rotated_Factors_thres[abs(Rotated_Factors_thres) < MIN_VALUE]<-NA
colnames(Rotated_Factors_thres)<- colnames(Rotated_Factors)
rownames(Rotated_Factors_thres)<- rownames(Rotated_Factors)

iprint.df(Rotated_Factors_thres, scale=TRUE)
```

## Step 6:  Save factor scores 

We can now either replace all initial variables used in this part with the factors scores or just select one of the initial variables for each of the selected factors in order to represent that factor. Here is how the factor scores  are for the first few respondents:

```{r}
NEW_ProjectData <- round(Rotated_Results$scores[,1:factors_selected,drop=F],2)
colnames(NEW_ProjectData)<-paste("DV (Factor)",1:ncol(NEW_ProjectData),sep=" ")

iprint.df(t(head(NEW_ProjectData, 10)), scale=TRUE)
```

### is this the number of dates??


# Part 2: Customer Segmentation 

```{r setupcluster, echo=TRUE, tidy=TRUE}
# Please ENTER then original raw attributes to use for the segmentation (the "segmentation attributes")
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
segmentation_attributes_used = c(12,10,11,8) #c(10,19,5,12,3) 

# Please ENTER then original raw attributes to use for the profiling of the segments (the "profiling attributes")
# Please use numbers, not column names, e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
profile_attributes_used = c(2:15) 

# Please ENTER the number of clusters to eventually use for this report
numb_clusters_used = 3 # for boats possibly use 5, for Mall_Visits use 3

# Please enter the method to use for the segmentation:
profile_with = "hclust" #  "hclust" or "kmeans"

# Please ENTER the distance metric eventually used for the clustering in case of hierarchical clustering 
# (e.g. "euclidean", "maximum", "manhattan", "canberra", "binary" or "minkowski" - see help(dist)). 
# DEFAULT is "euclidean"
distance_used = "euclidean"

# Please ENTER the hierarchical clustering method to use (options are:
# "ward", "single", "complete", "average", "mcquitty", "median" or "centroid").
# DEFAULT is "ward"
hclust_method = "ward.D"

# Please ENTER the kmeans clustering method to use (options are:
# "Hartigan-Wong", "Lloyd", "Forgy", "MacQueen").
# DEFAULT is "Lloyd"
kmeans_method = "Lloyd"

```


```{r}
# Same as the initial data
ProjectData <- ProjectData_INITIAL

segmentation_attributes_used <- intersect(segmentation_attributes_used, 1:ncol(ProjectData))
profile_attributes_used <- intersect(profile_attributes_used, 1:ncol(ProjectData))

ProjectData_segment <- ProjectData[,segmentation_attributes_used]
ProjectData_profile <- ProjectData[,profile_attributes_used]

ProjectData_scaled <- apply(ProjectData, 2, function(r) if (sd(r)!=0) (r-mean(r))/sd(r) else 0*r)
```


```{r}
euclidean_pairwise <- as.matrix(dist(head(ProjectData_segment, max_data_report), method="euclidean"))
euclidean_pairwise <- euclidean_pairwise*lower.tri(euclidean_pairwise) + euclidean_pairwise*diag(euclidean_pairwise) + 10e10*upper.tri(euclidean_pairwise)
euclidean_pairwise[euclidean_pairwise==10e10] <- NA
rownames(euclidean_pairwise) <- colnames(euclidean_pairwise) <- sprintf("Date.%02d", 1:max_data_report)

iprint.df(round(euclidean_pairwise))
```

## Step 5: Visualize Pair-wise Distances

We can see the histogram of, say, the first 2 variables (can you change the code chunk in the raw .Rmd file to see other variables?)

```{r}
variables_to_plot = 1:2
do.call(iplot.grid, lapply(variables_to_plot, function(n){
  iplot.hist(ProjectData_segment[, n], breaks=5, xlab = paste("Variable", n))
}))
```

```{r}
Pairwise_Distances <- dist(ProjectData_segment, method = distance_used) 
iplot.hist(Pairwise_Distances, breaks=10)
```

## Step 6: Method and Number of Segments

We need to select the clustering method to use, as well as the number of cluster. It may be useful to see the dendrogram from Hierarchical Clustering, to have a quick idea of how the data may be segmented and how many segments there may be. Here is the dendrogram for our data:

```{r}
Hierarchical_Cluster_distances <- dist(ProjectData_segment, method=distance_used)
Hierarchical_Cluster <- hclust(Hierarchical_Cluster_distances, method=hclust_method)
# Display dendogram
iplot.dendrogram(Hierarchical_Cluster)
# TODO: Draw dendogram with red borders around the 3 clusters
# rect.hclust(Hierarchical_Cluster, k=numb_clusters_used, border="red") 
```

We can also plot the "distances" traveled before we need to merge any of the lower and smaller in size clusters into larger ones - the heights of the tree branches that link the clusters as we traverse the tree from its leaves to its root. If we have n observations, this plot has n-1 numbers, we see the first 20 here. 
```{r}
num <- nrow(ProjectData) - 1
df1 <- cbind(as.data.frame(Hierarchical_Cluster$height[length(Hierarchical_Cluster$height):1]), c(1:num))
colnames(df1) <- c("distances","index")
iplot.df(melt(head(df1, 20), id="index"), xlab="Number of Components")
```

Here is the segment membership of the first `r max_data_report` respondents if we use hierarchical clustering:

```{r}
cluster_memberships_hclust <- as.vector(cutree(Hierarchical_Cluster, k=numb_clusters_used)) # cut tree into 3 clusters
cluster_ids_hclust=unique(cluster_memberships_hclust)

ProjectData_with_hclust_membership <- cbind(1:length(cluster_memberships_hclust),cluster_memberships_hclust)
colnames(ProjectData_with_hclust_membership)<-c("Observation Number","Cluster_Membership")

iprint.df(round(head(ProjectData_with_hclust_membership, max_data_report), 2))
```

while this is the segment membership if we use k-means:

```{r}
kmeans_clusters <- kmeans(ProjectData_segment,centers= numb_clusters_used, iter.max=2000, algorithm=kmeans_method)

ProjectData_with_kmeans_membership <- cbind(1:length(kmeans_clusters$cluster),kmeans_clusters$cluster)
colnames(ProjectData_with_kmeans_membership)<-c("Observation Number","Cluster_Membership")

iprint.df(round(head(ProjectData_with_kmeans_membership, max_data_report), 2))
```

## Step 7: Profile and interpret the segments 

In market segmentation one may use variables to **profile** the segments which are not the same (necessarily) as those used to **segment** the market: the latter may be, for example, attitude/needs related (you define segments based on what the customers "need"), while the former may be any information that allows a company to identify the defined customer segments (e.g. demographics, location, etc). Of course deciding which variables to use for segmentation and which to use for profiling (and then **activation** of the segmentation for business purposes) is largely subjective.  In this case we can use all survey questions for profiling for now - the `profile_attributes_used` variables selected below. 

There are many ways to do the profiling of the segments. For example, here we show how the *average* answers of the respondents *in each segment* compare to the *average answer of all respondents* using the ratio of the two.  The idea is that if in a segment the average response to a question is very different (e.g. away from ratio of 1) than the overall average, then that question may indicate something about the segment relative to the total population. 

Here are for example the profiles of the segments using the clusters found above.  First let's see just the average answer people gave to each question for the different segments as well as the total population:

```{r}
cluster_memberships_kmeans <- kmeans_clusters$cluster 
cluster_ids_kmeans <- unique(cluster_memberships_kmeans)

if (profile_with == "hclust"){
  cluster_memberships <- cluster_memberships_hclust
  cluster_ids <-  cluster_ids_hclust  
}
if (profile_with == "kmeans"){
  cluster_memberships <- cluster_memberships_kmeans
  cluster_ids <-  cluster_ids_kmeans
}

# WE WILL USE THESE IN THE CLASSIFICATION PART LATER
NewData = matrix(cluster_memberships,ncol=1)

population_average = matrix(apply(ProjectData_profile, 2, mean), ncol=1)
colnames(population_average) <- "Population"
Cluster_Profile_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_profile[(cluster_memberships==i), ], 2, mean))
if (ncol(ProjectData_profile) <2)
  Cluster_Profile_mean=t(Cluster_Profile_mean)
colnames(Cluster_Profile_mean) <- paste("Seg.", 1:length(cluster_ids), sep="")
cluster.profile <- cbind (population_average,Cluster_Profile_mean)

iprint.df(round(cluster.profile, 2))
```

We can also "visualize" the segments using **snake plots** for each cluster. For example, we can plot the means of the profiling variables for each of our clusters to better visualize differences between segments. For better visualization we plot the standardized profiling variables.

```{r}
ProjectData_scaled_profile = ProjectData_scaled[, profile_attributes_used,drop=F]

Cluster_Profile_standar_mean <- sapply(sort(cluster_ids), function(i) apply(ProjectData_scaled_profile[(cluster_memberships==i), ,drop = F], 2, mean))
if (ncol(ProjectData_scaled_profile) < 2)
  Cluster_Profile_standar_mean = t(Cluster_Profile_standar_mean)
colnames(Cluster_Profile_standar_mean) <- paste("Seg ", 1:length(cluster_ids), sep="")

iplot.df(melt(cbind.data.frame(idx=as.numeric(1:nrow(Cluster_Profile_standar_mean)), Cluster_Profile_standar_mean), id="idx"), xlab="Profiling variables (standardized)",  ylab="Mean of cluster")
```

We can also compare the averages of the profiling variables of each segment relative to the average of the variables across the whole population. This can also help us better understand whether  there are indeed clusters in our data (e.g. if all segments are much like the overall population, there may be no segments). For example, we can measure the ratios of the average for each cluster to the average of the population, minus 1, (e.g. `avg(cluster)` `/` `avg(population)` `-1`) for each segment and variable:

```{r}
population_average_matrix <- population_average[,"Population",drop=F] %*% matrix(rep(1,ncol(Cluster_Profile_mean)),nrow=1)
cluster_profile_ratios <- (ifelse(population_average_matrix==0, 0,Cluster_Profile_mean/population_average_matrix))
colnames(cluster_profile_ratios) <- paste("Seg.", 1:ncol(cluster_profile_ratios), sep="")
rownames(cluster_profile_ratios) <- colnames(ProjectData)[profile_attributes_used]
## printing the result in a clean-slate table
iprint.df(round(cluster_profile_ratios-1, 2))
```


## Step 8: Robustness Analysis

We should also consider the robustness of our analysis as we change the clustering method and parameters. Once we are comfortable with the solution we can finally answer our first business questions: 


# Part 3: Purchase Drivers

We will now use the [classification analysis ](http://inseaddataanalytics.github.io/INSEADAnalytics/CourseSessions/Sessions67/ClassificationAnalysisReading.html) methods to understand the key purchase drivers for boats (a similar analysis can be done for recommendation drivers). For simplicity we do not follow the "generic" steps of classification discussed in that reading, and only consider the classification and purchase drivers analysis for the segments we found above. 

We are interested in understanding the purchase drivers, hence our **dependent** variable is column 82 of the Boats data (`r colnames(ProjectData)[82]`) - why is that? We will use only the subquestions of **Question 16** of the case for now, and also select some of the parameters for this part of the analysis: 

```{r setupclassification, echo=TRUE, tidy=TRUE}
# Please ENTER the class (dependent) variable:
# Please use numbers, not column names! e.g. 82 uses the 82nd column are dependent variable.
# YOU NEED TO MAKE SURE THAT THE DEPENDENT VARIABLES TAKES ONLY 2 VALUES: 0 and 1!!!
dependent_variable= 2

# Please ENTER the attributes to use as independent variables 
# Please use numbers, not column names! e.g. c(1:5, 7, 8) uses columns 1,2,3,4,5,7,8
independent_variables= c(5) # use 54-80 for boats

# Please ENTER the profit/cost values for the correctly and wrong classified data:
actual_1_predict_1 = 100
actual_1_predict_0 = -75
actual_0_predict_1 = -50
actual_0_predict_0 = 0

# Please ENTER the probability threshold above which an observations  
# is predicted as class 1:
Probability_Threshold=50 # between 1 and 99%

# Please ENTER the percentage of data used for estimation
estimation_data_percent = 80
validation_data_percent = 10

# Please enter 0 if you want to "randomly" split the data in estimation and validation/test
random_sampling = 0

# Tree parameter
# PLEASE ENTER THE Tree (CART) complexity control cp (e.g. 0.001 to 0.02, depending on the data)
CART_cp = 0.01

# Please enter the minimum size of a segment for the analysis to be done only for that segment
min_segment = 100
```

```{r}
ProjectData = ProjectData_INITIAL # Just to initialize the data

Probability_Threshold = Probability_Threshold/100 # make it between 0 and 1

dependent_variable = unique(sapply(dependent_variable,function(i) min(ncol(ProjectData), max(i,1))))
independent_variables = unique(sapply(independent_variables,function(i) min(ncol(ProjectData), max(i,1))))

if (length(unique(ProjectData[,dependent_variable])) !=2){
  cat("\n*****\n BE CAREFUL, THE DEPENDENT VARIABLE TAKES MORE THAN 2 VALUES...")
  cat("\nSplitting it around its median...\n*****\n ")
  new_dependent = ProjectData[,dependent_variable] >= median(ProjectData[,dependent_variable])
  ProjectData[,dependent_variable] <- 1*new_dependent
}

Profit_Matrix = matrix(c(actual_1_predict_1, actual_0_predict_1, actual_1_predict_0, actual_0_predict_0), ncol=2)
colnames(Profit_Matrix)<- c("Predict 1", "Predict 0")
rownames(Profit_Matrix) <- c("Actual 1", "Actual 0")
test_data_percent = 100-estimation_data_percent-validation_data_percent
CART_control = rpart.control(cp = CART_cp)
```


We will use two classification trees and logistic regression. You can select "complexity" control for one of the classification trees in the code chunk of the raw .Rmd file here

```{r CART_control, echo=TRUE, tidy=TRUE}
CART_control = 0.001
```

```{r}
# FIrst we split the data in estimation, validation, and test

if (random_sampling){
  estimation_data_ids=sample.int(nrow(ProjectData),floor(estimation_data_percent*nrow(ProjectData)/100))
  non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
  validation_data_ids=non_estimation_data[sample.int(length(non_estimation_data), floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))]
  } else {
    estimation_data_ids=1:floor(estimation_data_percent*nrow(ProjectData)/100)
    non_estimation_data = setdiff(1:nrow(ProjectData),estimation_data_ids)
    validation_data_ids = (tail(estimation_data_ids,1)+1):(tail(estimation_data_ids,1) + floor(validation_data_percent/(validation_data_percent+test_data_percent)*length(non_estimation_data)))
    }

test_data_ids = setdiff(1:nrow(ProjectData), union(estimation_data_ids,validation_data_ids))

estimation_data=ProjectData[estimation_data_ids,]
validation_data=ProjectData[validation_data_ids,]
test_data=ProjectData[test_data_ids,]
```

```{r}
# just name the variables numerically so that they look ok on the tree plots
independent_variables_nolabel = paste("IV", 1:length(independent_variables), sep="")

estimation_data_nolabel = cbind(estimation_data[,dependent_variable], estimation_data[,independent_variables])
colnames(estimation_data_nolabel)<- c(colnames(estimation_data)[dependent_variable],independent_variables_nolabel)

validation_data_nolabel = cbind(validation_data[,dependent_variable], validation_data[,independent_variables])
colnames(validation_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

test_data_nolabel = cbind(test_data[,dependent_variable], test_data[,independent_variables])
colnames(test_data_nolabel)<- c(dependent_variable,independent_variables_nolabel)

estimation_data_nolabel = data.frame(estimation_data_nolabel)
validation_data_nolabel = data.frame(validation_data_nolabel)
test_data_nolabel = data.frame(test_data_nolabel)

estimation_data = data.frame(estimation_data)
validation_data = data.frame(validation_data)
test_data = data.frame(test_data)
```

```{r}
formula=paste(colnames(estimation_data)[dependent_variable],paste(Reduce(paste,sapply(head(independent_variables_nolabel,-1), function(i) paste(i,"+",sep=""))),tail(independent_variables_nolabel,1),sep=""),sep="~")
CART_tree<-rpart(formula, data= estimation_data_nolabel,method="class", control=CART_control)

rpart.plot(CART_tree, box.palette="OrBu", type=3, extra=1, fallen.leaves=F, branch.lty=3)
```


